Multi-Agent Pathfinding with Matching using Increasing Cost Tree Search

### 2021-06-27 docs/paper.tex
Finished Section 5,6

	\begin{figure}[H]
		\centering
		\begin{subfigure}{0.40\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/counter-example-ep-2}
			\caption{Counter-example instance.}
			\label{fig:counter-example:a}
		\end{subfigure}
		\begin{subfigure}{0.49\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/counter-example-tree}
			\caption{Pruned ICT that is traversed.}
			\label{fig:counter-example:b}
		\end{subfigure}
		\caption{Counter-example for enhanced pruning combined with child pruning. If the black and green agent pair is considered last, an ICT is traversed while solving the instance that does not contain an optimal solution.}
		\label{fig:counter-example}
	\end{figure}
	\begin{proof}
		Consider Figure \ref{fig:counter-example:a} with a red agent $r$, green agent $g$ and black agent $b$. It can be seen that $(3,8,5)$ with SIC $16$ is an optimal solution for this instance, which corresponds to $r$ moving aside, allowing $g$ to pass.  Say enhanced pruning with pairs is applied to these three agents at $(c^*_r,c^*_g,c^*_b) = (0,8,5)$. Let the enhanced pruning order be $(r,g)$, $(r,b)$, $(g,b)$ without loss of generality. $(r,g)$ is solvable but will leave for $g$ out of the two paths of length $8$ only the clock-wise path. $(r,b)$ will not change either MDD in this node. When $(g,b)$ is evaluated, it is found that the remaining clockwise path for $g$ and $b$'s path are incompatible (edge conflict) so $(0,9,5)$ and $(0,8,6)$ are generated. By repeating this order of enhanced pruning with child pruning, the ICT shown in Figure \ref{fig:counter-example:b} is traversed. None of the nodes with SIC 16 in this ICT is a solution, meaning that enhanced pruning combined with child pruning will return a suboptimal solution to the instance depicted in Figure \ref{fig:counter-example:a}, making this pruning strategy suboptimal.
	\end{proof}

@@

	CBM solving all problems with one team (Figures \ref{fig:r-25-1-p},\ref{fig:r-75-1-p}), problems equivalent to anonymous MAPF instances, is at first surprising but follows directly from the definition of the CBM method. CBM solves single teams as anonymous MAPF instances in polynomial time, meaning that solving scenarios with a single team reduces to a polynomial-time procedure, namely SSP. Following the results reported in \cite{ma2016}, it is expected that as the number of teams approaches $k$, CBM performance will deteriorate and thus become more similar to the results for exhaustive methods here reported.
	%	Still, some implementations might be more optimized than others so this might be a factor in performance characteristics and in particular in relative performance differences.
	\section{Conclusions and Future work}
	\label{conclusions}
	\label{future}
	The goal of this research was to seek methods for solving MAPFM based on ICTS. In this, the research has succeeded, but the scalability of the algorithms leaves something to be desired. Exhaustive methods, like those discussed in detail in Subsection \ref{exhaustive}, cannot be expected to scale well as teams grow larger due to their complexity, no matter what optimizations are put in place. Nonetheless, as was shown in Section \ref{experiments}, the more advanced exhaustive ICTS algorithms consistently outperform the ICTS-m based algorithms. Compared to the SIC-adaptation of the CBM method \cite{baauw2021}, it was seen that exhaustive methods generally perform relatively poorly if there are fewer teams and fewer obstacles, while being competitive once teams grow smaller and maps become more constrained. Delineating under which conditions exhaustive methods outperform CBM and vice versa is something for further research.

@@

	A related observation is that the algorithms introduced in this work do not rely on the property of agents and goals having one color. Both exhaustive ICTS and ICTS-m could easily be adapted to support one or more colors (teams) for each agent and goal while still guaranteeing optimality, in contrast to CBM \cite{ma2016} and its SIC adaptation \cite{baauw2021} which specifically use this property to plan single teams in polynomial time. Future research could study the multi-color MAPFM generalization and the application of methods described in this work to that problem.

>>>>

	\begin{figure}[H]
		\centering
		\begin{subfigure}{0.44\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/counter-example-ep-2}
			\caption{Counter-example instance.}
			\label{fig:counter-example:a}
		\end{subfigure}
		\begin{subfigure}{0.49\textwidth}
			\centering
			\includegraphics[width=\linewidth]{img/counter-example-tree}
			\caption{Pruned ICT that is traversed.}
			\label{fig:counter-example:b}
		\end{subfigure}
		\caption{Counter-example for enhanced pruning combined with child pruning. If the black and green agent pair is considered last, an ICT is traversed while solving the instance that does not contain an optimal solution.}
		\label{fig:counter-example}
	\end{figure}
	\begin{proof}
		Consider Figure \ref{fig:counter-example:a} with a red agent $r$, green agent $g$ and black agent $b$. It can be seen that $(3,8,5)$ with SIC $16$ is an optimal solution for this instance, which corresponds to $r$ moving aside, allowing $g$ to pass.  Say enhanced pruning with pairs is applied to these three agents at $(c^*_r,c^*_g,c^*_b) = (0,8,5)$. Let the enhanced pruning order be $(r,g)$, $(r,b)$, $(g,b)$ without loss of generality. $(r,g)$ is solvable but will leave for $g$ out of the two paths of length $8$ only the clock-wise path. $(r,b)$ will not change either MDD in this node. When $(g,b)$ is evaluated, it is found that the remaining clockwise path for $g$ and $b$'s path are incompatible (edge conflict) so $(0,9,5)$ and $(0,8,6)$ are generated. By repeating this order of enhanced pruning with child pruning, the ICT shown in Figure \ref{fig:counter-example:b} is traversed. None of the nodes with SIC 16 in this ICT is a solution, meaning that enhanced pruning combined with child pruning will return a suboptimal solution to the instance depicted in Figure \ref{fig:counter-example:a}, making this pruning strategy suboptimal.
	\end{proof}

@@

	CBM solving all problems with one team (Figures \ref{fig:r-25-1-p},\ref{fig:r-75-1-p}) is at first surprising but follows directly from the definition of the CBM method. CBM solves single teams as anonymous MAPF instances in polynomial time, meaning that solving scenarios with a single team reduces to a polynomial-time procedure, namely SSP. Following the results reported in \cite{ma2016}, it is expected that as the number of teams approaches $k$, CBM performance will deteriorate and thus become more similar to the results for exhaustive methods here reported.
	%	Still, some implementations might be more optimized than others so this might be a factor in performance characteristics and in particular in relative performance differences.
	\section{Conclusions and Future work}
	\label{conclusions}
	\label{future}
	The goal of this research was to seek methods for solving MAPFM based on ICTS. In this, the research has succeeded, but the scalability of the algorithms leaves something to be desired. Exhaustive methods, like those discussed in detail in Subsection \ref{exhaustive}, cannot be expected to scale well as teams grow larger due to their complexity, no matter what optimizations are put in place. Nonetheless, as was shown in Section \ref{experiments}, the more advanced exhaustive ICTS algorithms consistently outperform the ICTS-m based algorithms. Compared to the SIC-adaptation of the CBM method \cite{baauw2021}, it was seen that exhaustive methods perform relatively poorly if there are fewer teams and fewer obstacles, while being competitive once teams grow smaller and maps become more constrained. Delineating under which conditions exhaustive methods outperform CBM and vice versa is something for further research.

@@

	A related observation is that the algorithms introduced in this work in their current form do not rely on the property of agents and goals having one color (team). Both exhaustive ICTS and ICTS-m could easily be adapted to support one or more colors for each agent and goal while still guaranteeing optimality, in contrast to CBM \cite{ma2016} and its SIC adaptation \cite{baauw2021} which specifically use this property to plan single teams in polynomial time. Future research could study the multi-color MAPFM generalization and the application of methods described in this work to that problem.
